---
title: "Experimental evolution of TPCs: Cold-tolerant species evolve larger fitness gains at higher temperatures but do not reach the warm-tolerant species"
author: "Jennifer Molinet"
date: "2025-02-14"
output: html_document
---

To test if and how TPCs evolve across the yeast genus, we performed an adaptive evolution experiment in which we gradually increased temperature from 25 °C to 40 °C every 50 generations or until strains went extinct.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Libraries**

```{r}
suppressMessages(library(tidyverse))
library(ggplot2)
library(grid)
suppressMessages(library(viridis))
library(patchwork)
library(hrbrthemes)
suppressMessages(library(igraph))
library(ggraph)
library(colormap)
library(DiagrammeR)
library(scales)
library(ggbreak)
library(reshape2)
suppressMessages(library(lme4))
suppressMessages(library(cowplot))
```

**Directories**

```{r}
output_dir <- "../results/Section2/"
input_dir <- "../data/Section2"
```

Overall, cold-tolerant species did not survive beyond 34 °C, whereas warm-tolerant species reached temperatures between 34 and 40 °C, with both S. cerevisiae strains surviving the maximum temperature of 40 °C after 600 generations.

Plotting figure 3A

```{r figure3A, fig.width=12, fig.height=6}
ev2 <- ggplot()+
  #SE2
  geom_rect(aes(xmin=0,xmax=1,ymin=0,ymax=1),
            fill="#ffe5e5", alpha=1,color="black")+ #25
  geom_rect(aes(xmin=1,xmax=2,ymin=0,ymax=1),
            fill="#ffcccc", alpha=1,color="black")+ #28
  geom_rect(aes(xmin=2,xmax=3,ymin=0,ymax=1),
            fill="#ffb2b2", alpha=1,color="black")+ #29
  geom_rect(aes(xmin=3,xmax=4,ymin=0,ymax=1),
            fill="#ff9999", alpha=1,color="black")+ #30
  geom_rect(aes(xmin=4,xmax=5,ymin=0,ymax=1),
            fill="#ff7f7f", alpha=1,color="black")+ #31
  #SE1
  geom_rect(aes(xmin=0,xmax=1,ymin=1,ymax=2),
            fill="#ffe5e5", alpha=1,color="black")+ #25
  geom_rect(aes(xmin=1,xmax=2,ymin=1,ymax=2),
            fill="#ffcccc", alpha=1,color="black")+ #28
  geom_rect(aes(xmin=4,xmax=5,ymin=1,ymax=2),
            fill="#ff7f7f", alpha=1,color="black")+ #31
  geom_rect(aes(xmin=5,xmax=6,ymin=1,ymax=2),
            fill="#ff6666", alpha=1,color="black")+ #32
  geom_rect(aes(xmin=6,xmax=7,ymin=1,ymax=2),
            fill="#ff4c4c", alpha=1,color="black")+ #33
  geom_rect(aes(xmin=7,xmax=8,ymin=1,ymax=2),
            fill="#ff3232", alpha=1,color="black")+ #34
  #SU1
  geom_rect(aes(xmin=0,xmax=1,ymin=2,ymax=3),
            fill="#ffe5e5", alpha=1,color="black")+ #25
  geom_rect(aes(xmin=1,xmax=2,ymin=2,ymax=3),
            fill="#ffcccc", alpha=1,color="black")+ #28
  geom_rect(aes(xmin=4,xmax=5,ymin=2,ymax=3),
            fill="#ff7f7f", alpha=1,color="black")+ #31
  geom_rect(aes(xmin=5,xmax=6,ymin=2,ymax=3),
            fill="#ff6666", alpha=1,color="black")+ #32
  geom_rect(aes(xmin=6,xmax=7,ymin=2,ymax=3),
            fill="#ff4c4c", alpha=1,color="black")+ #33
  #SU2
  geom_rect(aes(xmin=0,xmax=1,ymin=3,ymax=4),
            fill="#ffe5e5", alpha=1,color="black")+ #25
  geom_rect(aes(xmin=1,xmax=2,ymin=3,ymax=4),
            fill="#ffcccc", alpha=1,color="black")+ #28
  geom_rect(aes(xmin=4,xmax=5,ymin=3,ymax=4),
            fill="#ff7f7f", alpha=1,color="black")+ #31
  geom_rect(aes(xmin=5,xmax=6,ymin=3,ymax=4),
            fill="#ff6666", alpha=1,color="black")+ #32
  geom_rect(aes(xmin=6,xmax=7,ymin=3,ymax=4),
            fill="#ff4c4c", alpha=1,color="black")+ #33
  #SA2
  geom_rect(aes(xmin=0,xmax=1,ymin=4,ymax=5),
            fill="#ffe5e5", alpha=1,color="black")+ #25
  geom_rect(aes(xmin=1,xmax=2,ymin=4,ymax=5),
            fill="#ffcccc", alpha=1,color="black")+ #28
  geom_rect(aes(xmin=4,xmax=5,ymin=4,ymax=5),
            fill="#ff7f7f", alpha=1,color="black")+ #31
  geom_rect(aes(xmin=5,xmax=6,ymin=4,ymax=5),
            fill="#ff6666", alpha=1,color="black")+ #32
  geom_rect(aes(xmin=6,xmax=7,ymin=4,ymax=5),
            fill="#ff4c4c", alpha=1,color="black")+ #33
  #SK1
  geom_rect(aes(xmin=0,xmax=1,ymin=5,ymax=6),
            fill="#ffe5e5", alpha=1,color="black")+ #25
  geom_rect(aes(xmin=1,xmax=2,ymin=5,ymax=6),
            fill="#ffcccc", alpha=1,color="black")+ #28
  geom_rect(aes(xmin=4,xmax=5,ymin=5,ymax=6),
            fill="#ff7f7f", alpha=1,color="black")+ #31
  geom_rect(aes(xmin=5,xmax=6,ymin=5,ymax=6),
            fill="#ff6666", alpha=1,color="black")+ #32
  #SK2
  geom_rect(aes(xmin=0,xmax=1,ymin=6,ymax=7),
            fill="#ffe5e5", alpha=1,color="black")+ #25
  geom_rect(aes(xmin=1,xmax=2,ymin=6,ymax=7),
            fill="#ffcccc", alpha=1,color="black")+ #28
  geom_rect(aes(xmin=4,xmax=5,ymin=6,ymax=7),
            fill="#ff7f7f", alpha=1,color="black")+ #31
  geom_rect(aes(xmin=5,xmax=6,ymin=6,ymax=7),
            fill="#ff6666", alpha=1,color="black")+ #32
  #SJ1
  geom_rect(aes(xmin=0,xmax=1,ymin=7,ymax=8),
            fill="#ffe5e5", alpha=1,color="black")+ #25
  geom_rect(aes(xmin=1,xmax=2,ymin=7,ymax=8),
            fill="#ffcccc", alpha=1,color="black")+ #28
  geom_rect(aes(xmin=4,xmax=5,ymin=7,ymax=8),
            fill="#ff7f7f", alpha=1,color="black")+ #31
  geom_rect(aes(xmin=5,xmax=6,ymin=7,ymax=8),
            fill="#ff6666", alpha=1,color="black")+ #32
  geom_rect(aes(xmin=6,xmax=7,ymin=7,ymax=8),
            fill="#ff4c4c", alpha=1,color="black")+ #33
  geom_rect(aes(xmin=7,xmax=8,ymin=7,ymax=8),
            fill="#ff3232", alpha=1,color="black")+ #34
  geom_rect(aes(xmin=8,xmax=9,ymin=7,ymax=8),
            fill="#ff1919", alpha=1,color="black")+ #35
  #SJ2
  geom_rect(aes(xmin=0,xmax=1,ymin=8,ymax=9),
            fill="#ffe5e5", alpha=1,color="black")+ #25
  geom_rect(aes(xmin=1,xmax=2,ymin=8,ymax=9),
            fill="#ffcccc", alpha=1,color="black")+ #28
  geom_rect(aes(xmin=4,xmax=5,ymin=8,ymax=9),
            fill="#ff7f7f", alpha=1,color="black")+ #31
  geom_rect(aes(xmin=5,xmax=6,ymin=8,ymax=9),
            fill="#ff6666", alpha=1,color="black")+ #32
  geom_rect(aes(xmin=6,xmax=7,ymin=8,ymax=9),
            fill="#ff4c4c", alpha=1,color="black")+ #33
  geom_rect(aes(xmin=7,xmax=8,ymin=8,ymax=9),
            fill="#ff3232", alpha=1,color="black")+ #34
  geom_rect(aes(xmin=8,xmax=9,ymin=8,ymax=9),
            fill="#ff1919", alpha=1,color="black")+ #35
  #SM1
  geom_rect(aes(xmin=0,xmax=1,ymin=9,ymax=10),
            fill="#ffe5e5", alpha=1,color="black")+ #25
  geom_rect(aes(xmin=1,xmax=2,ymin=9,ymax=10),
            fill="#ffcccc", alpha=1,color="black")+ #28
  geom_rect(aes(xmin=4,xmax=5,ymin=9,ymax=10),
            fill="#ff7f7f", alpha=1,color="black")+ #31
  geom_rect(aes(xmin=5,xmax=6,ymin=9,ymax=10),
            fill="#ff6666", alpha=1,color="black")+ #32
  geom_rect(aes(xmin=6,xmax=7,ymin=9,ymax=10),
            fill="#ff4c4c", alpha=1,color="black")+ #33
  geom_rect(aes(xmin=7,xmax=8,ymin=9,ymax=10),
            fill="#ff3232", alpha=1,color="black")+ #34
  #SM2
  geom_rect(aes(xmin=0,xmax=1,ymin=10,ymax=11),
            fill="#ffe5e5", alpha=1,color="black")+ #25
  geom_rect(aes(xmin=1,xmax=2,ymin=10,ymax=11),
            fill="#ffcccc", alpha=1,color="black")+ #28
  geom_rect(aes(xmin=4,xmax=5,ymin=10,ymax=11),
            fill="#ff7f7f", alpha=1,color="black")+ #31
  geom_rect(aes(xmin=5,xmax=6,ymin=10,ymax=11),
            fill="#ff6666", alpha=1,color="black")+ #32
  geom_rect(aes(xmin=6,xmax=7,ymin=10,ymax=11),
            fill="#ff4c4c", alpha=1,color="black")+ #33
  geom_rect(aes(xmin=7,xmax=8,ymin=10,ymax=11),
            fill="#ff3232", alpha=1,color="black")+ #34
  geom_rect(aes(xmin=8,xmax=9,ymin=10,ymax=11),
            fill="#ff1919", alpha=1,color="black")+ #35
  geom_rect(aes(xmin=9,xmax=10,ymin=10,ymax=11),
            fill="#ff0000", alpha=1,color="black")+ #36
  #SP1
  geom_rect(aes(xmin=0,xmax=1,ymin=11,ymax=12),
            fill="#ffe5e5", alpha=1,color="black")+ #25
  geom_rect(aes(xmin=1,xmax=2,ymin=11,ymax=12),
            fill="#ffcccc", alpha=1,color="black")+ #28
  geom_rect(aes(xmin=4,xmax=5,ymin=11,ymax=12),
            fill="#ff7f7f", alpha=1,color="black")+ #31
  geom_rect(aes(xmin=5,xmax=6,ymin=11,ymax=12),
            fill="#ff6666", alpha=1,color="black")+ #32
  geom_rect(aes(xmin=6,xmax=7,ymin=11,ymax=12),
            fill="#ff4c4c", alpha=1,color="black")+ #33
  geom_rect(aes(xmin=7,xmax=8,ymin=11,ymax=12),
            fill="#ff3232", alpha=1,color="black")+ #34
  geom_rect(aes(xmin=8,xmax=9,ymin=11,ymax=12),
            fill="#ff1919", alpha=1,color="black")+ #35
  #SP2
  geom_rect(aes(xmin=0,xmax=1,ymin=12,ymax=13),
            fill="#ffe5e5", alpha=1,color="black")+ #25
  geom_rect(aes(xmin=1,xmax=2,ymin=12,ymax=13),
            fill="#ffcccc", alpha=1,color="black")+ #28
  geom_rect(aes(xmin=4,xmax=5,ymin=12,ymax=13),
            fill="#ff7f7f", alpha=1,color="black")+ #31
  geom_rect(aes(xmin=5,xmax=6,ymin=12,ymax=13),
            fill="#ff6666", alpha=1,color="black")+ #32
  geom_rect(aes(xmin=6,xmax=7,ymin=12,ymax=13),
            fill="#ff4c4c", alpha=1,color="black")+ #33
  geom_rect(aes(xmin=7,xmax=8,ymin=12,ymax=13),
            fill="#ff3232", alpha=1,color="black")+ #34
  geom_rect(aes(xmin=8,xmax=9,ymin=12,ymax=13),
            fill="#ff1919", alpha=1,color="black")+ #35
  geom_rect(aes(xmin=9,xmax=10,ymin=12,ymax=13),
            fill="#ff0000", alpha=1,color="black")+ #36
  #SC
  geom_rect(aes(xmin=0,xmax=1,ymin=13,ymax=14),
            fill="#ffe5e5", alpha=1,color="black")+ #25
  geom_rect(aes(xmin=1,xmax=2,ymin=13,ymax=14),
            fill="#ffcccc", alpha=1,color="black")+ #28
  geom_rect(aes(xmin=4,xmax=5,ymin=13,ymax=14),
            fill="#ff7f7f", alpha=1,color="black")+ #31
  geom_rect(aes(xmin=5,xmax=6,ymin=13,ymax=14),
            fill="#ff6666", alpha=1,color="black")+ #32
  geom_rect(aes(xmin=6,xmax=7,ymin=13,ymax=14),
            fill="#ff4c4c", alpha=1,color="black")+ #33
  geom_rect(aes(xmin=7,xmax=8,ymin=13,ymax=14),
            fill="#ff3232", alpha=1,color="black")+ #34
  geom_rect(aes(xmin=8,xmax=9,ymin=13,ymax=14),
            fill="#ff1919", alpha=1,color="black")+ #35
  geom_rect(aes(xmin=9,xmax=10,ymin=13,ymax=14),
            fill="#ff0000", alpha=1,color="black")+ #36
  #SC1
  geom_rect(aes(xmin=0,xmax=1,ymin=14,ymax=15),
            fill="#ffe5e5", alpha=1,color="black")+ #25
  geom_rect(aes(xmin=1,xmax=2,ymin=14,ymax=15),
            fill="#ffcccc", alpha=1,color="black")+ #28
  geom_rect(aes(xmin=4,xmax=5,ymin=14,ymax=15),
            fill="#ff7f7f", alpha=1,color="black")+ #31
  geom_rect(aes(xmin=5,xmax=6,ymin=14,ymax=15),
            fill="#ff6666", alpha=1,color="black")+ #32
  geom_rect(aes(xmin=6,xmax=7,ymin=14,ymax=15),
            fill="#ff4c4c", alpha=1,color="black")+ #33
  geom_rect(aes(xmin=7,xmax=8,ymin=14,ymax=15),
            fill="#ff3232", alpha=1,color="black")+ #34
  geom_rect(aes(xmin=8,xmax=9,ymin=14,ymax=15),
            fill="#ff1919", alpha=1,color="black")+ #35
  geom_rect(aes(xmin=9,xmax=10,ymin=14,ymax=15),
            fill="#ff0000", alpha=1,color="black")+ #36
  geom_rect(aes(xmin=10,xmax=11,ymin=14,ymax=15),
            fill="#e50000", alpha=1,color="black")+ #37
  geom_rect(aes(xmin=11,xmax=12,ymin=14,ymax=15),
            fill="#cc0000", alpha=1,color="black")+ #38
  geom_rect(aes(xmin=12,xmax=13,ymin=14,ymax=15),
            fill="#b20000", alpha=1,color="black")+ #39
  geom_rect(aes(xmin=13,xmax=14,ymin=14,ymax=15),
            fill="#990000", alpha=1,color="black")+ #40
  #SC2
  geom_rect(aes(xmin=0,xmax=1,ymin=15,ymax=16),
            fill="#ffe5e5", alpha=1,color="black")+ #25
  geom_rect(aes(xmin=1,xmax=2,ymin=15,ymax=16),
            fill="#ffcccc", alpha=1,color="black")+ #28
  geom_rect(aes(xmin=4,xmax=5,ymin=15,ymax=16),
            fill="#ff7f7f", alpha=1,color="black")+ #31
  geom_rect(aes(xmin=5,xmax=6,ymin=15,ymax=16),
            fill="#ff6666", alpha=1,color="black")+ #32
  geom_rect(aes(xmin=6,xmax=7,ymin=15,ymax=16),
            fill="#ff4c4c", alpha=1,color="black")+ #33
  geom_rect(aes(xmin=7,xmax=8,ymin=15,ymax=16),
            fill="#ff3232", alpha=1,color="black")+ #34
  geom_rect(aes(xmin=8,xmax=9,ymin=15,ymax=16),
            fill="#ff1919", alpha=1,color="black")+ #35
  geom_rect(aes(xmin=9,xmax=10,ymin=15,ymax=16),
            fill="#ff0000", alpha=1,color="black")+ #36
  geom_rect(aes(xmin=10,xmax=11,ymin=15,ymax=16),
            fill="#e50000", alpha=1,color="black")+ #37
  geom_rect(aes(xmin=11,xmax=12,ymin=15,ymax=16),
            fill="#cc0000", alpha=1,color="black")+ #38
  geom_rect(aes(xmin=12,xmax=13,ymin=15,ymax=16),
            fill="#b20000", alpha=1,color="black")+ #39
  geom_rect(aes(xmin=13,xmax=14,ymin=15,ymax=16),
            fill="#990000", alpha=1,color="black")+ #40
  # Strain label
  annotate("text",x=-0.5,y=0.5,label="Se-6",fontface="bold",size=4)+
  annotate("text",x=-0.5,y=1.5,label="Se-1",fontface="bold",size=4)+
  annotate("text",x=-0.5,y=2.5,label="Su-4",fontface="bold",size=4)+
  annotate("text",x=-0.5,y=3.5,label="Su-2",fontface="bold",size=4)+
  annotate("text",x=-0.5,y=4.5,label="Sa-4",fontface="bold",size=4)+
  annotate("text",x=-0.5,y=5.5,label="Sk-1",fontface="bold",size=4)+
  annotate("text",x=-0.5,y=6.5,label="Sk-4",fontface="bold",size=4)+
  annotate("text",x=-0.5,y=7.5,label="Sj-2",fontface="bold",size=4)+
  annotate("text",x=-0.5,y=8.5,label="Sj-1",fontface="bold",size=4)+
  annotate("text",x=-0.5,y=9.5,label="Sm-1",fontface="bold",size=4)+
  annotate("text",x=-0.5,y=10.5,label="Sm-6",fontface="bold",size=4)+
  annotate("text",x=-0.5,y=11.5,label="Sp-5",fontface="bold",size=4)+
  annotate("text",x=-0.5,y=12.5,label="Sp-8",fontface="bold",size=4)+
  annotate("text",x=-0.5,y=13.5,label="Sc-2",fontface="bold",size=4)+
  annotate("text",x=-0.5,y=14.5,label="Sc-8",fontface="bold",size=4)+
  annotate("text",x=-0.5,y=15.5,label="Sc-13",fontface="bold",size=4)+
  # Temperature
  annotate("text",x=0.5,y=-0.5,label="25°C",fontface="bold",size=4)+
  annotate("text",x=1.5,y=-0.5,label="28°C",fontface="bold",size=4)+
  annotate("text",x=2.5,y=-0.5,label="29°C",fontface="bold",size=4)+
  annotate("text",x=3.5,y=-0.5,label="30°C",fontface="bold",size=4)+
  annotate("text",x=4.5,y=-0.5,label="31°C",fontface="bold",size=4)+
  annotate("text",x=5.5,y=-0.5,label="32°C",fontface="bold",size=4)+
  annotate("text",x=6.5,y=-0.5,label="33°C",fontface="bold",size=4)+
  annotate("text",x=7.5,y=-0.5,label="34°C",fontface="bold",size=4)+
  annotate("text",x=8.5,y=-0.5,label="35°C",fontface="bold",size=4)+
  annotate("text",x=9.5,y=-0.5,label="36°C",fontface="bold",size=4)+
  annotate("text",x=10.5,y=-0.5,label="37°C",fontface="bold",size=4)+
  annotate("text",x=11.5,y=-0.5,label="38°C",fontface="bold",size=4)+
  annotate("text",x=12.5,y=-0.5,label="39°C",fontface="bold",size=4)+
  annotate("text",x=13.5,y=-0.5,label="40°C",fontface="bold",size=4)+
  # Generations
  annotate("text",x=5.5,y=0.5,label="250 gen",fontface="bold",size=3)+
  annotate("text",x=8.5,y=1.5,label="350 gen",fontface="bold",size=3)+
  annotate("text",x=7.5,y=2.5,label="250 gen",fontface="bold",size=3)+
  annotate("text",x=7.5,y=3.5,label="250 gen",fontface="bold",size=3)+
  annotate("text",x=7.5,y=4.5,label="250 gen",fontface="bold",size=3)+
  annotate("text",x=6.5,y=5.5,label="200 gen",fontface="bold",size=3)+
  annotate("text",x=6.5,y=6.5,label="250 gen",fontface="bold",size=3)+
  annotate("text",x=9.5,y=7.5,label="350 gen",fontface="bold",size=3)+
  annotate("text",x=9.5,y=8.5,label="350 gen",fontface="bold",size=3)+
  annotate("text",x=8.5,y=9.5,label="300 gen",fontface="bold",size=3)+
  annotate("text",x=10.5,y=10.5,label="400 gen",fontface="bold",size=3)+
  annotate("text",x=9.5,y=11.5,label="350 gen",fontface="bold",size=3)+
  annotate("text",x=10.5,y=12.5,label="400 gen",fontface="bold",size=3)+
  annotate("text",x=10.5,y=13.5,label="400 gen",fontface="bold",size=3)+
  annotate("text",x=14.5,y=14.5,label="600 gen",fontface="bold",size=3)+
  annotate("text",x=14.5,y=15.5,label="600 gen",fontface="bold",size=3)+
  theme_bw() +
  labs(x ='Temperature', y="Strain")+
  theme(axis.ticks = element_blank(), axis.text.y = element_blank(),
        axis.text.x = element_blank(),
        axis.title.x = element_text(size = 14,face="bold"),
        axis.title.y = element_text( size = 14,face="bold"))+
  scale_y_continuous(breaks = seq(2,16,2))+
  scale_x_continuous(breaks = seq(0,14,2))

ev2
ggsave(filename = file.path(output_dir, "Figure_3A.tiff"), plot = ev2, width = 12, height = 6, dpi = 300)
```


**Loading data**
```{r}
file_path <- file.path(input_dir, "Kinetic_parameters.txt")
d <- read.table(file_path, header = TRUE, stringsAsFactors = FALSE)
d$Species <- ordered(d$Species, levels = c("SE", "SU", "SA", "SK", "SJ","SM","SP","SC"))
head(d)
```

**Determining relative fitness between evolved population and ancestral strains**

We then compared the evolution of relative fitness (difference in maximum growth rate between evolved populations and their ancestors) of the populations evolved under constant (25 °C, condition C) and increasing temperature (condition I) 

Plotting figure 3B

```{r}
d1 <- d %>%
  dplyr::group_by(Strain,Strain_2,Condition,Line,Generation,Temperature,Highlight,Species, Line.1) %>%
  dplyr::summarise(sd_rate = sd(Rate_fitness),
                   rate = mean(Rate_fitness),
                   sd_OD = sd(OD_fitness),
                   OD = mean(OD_fitness)) %>%
  ungroup()
d1 <- as.data.frame(d1)

d2 <- d1 %>% select(Strain,Strain_2,Condition,Line,Generation,Temperature,Highlight,Species,Line.1,rate) %>%
  dcast(Strain+Strain_2+Condition+Line+Generation+Highlight+Species+Line.1~Temperature, value.var = "rate")

d3 <- d1 %>% select(Strain,Strain_2,Condition,Line,Generation,Temperature,Highlight,Species,Line.1,sd_rate) %>%
  dcast(Strain+Strain_2+Condition+Line+Generation+Highlight+Species+Line.1~Temperature, value.var = "sd_rate")

Temp_name <- c("Strain","Strain_number","Condition","Line","Generation","Highlight","Species","Line number" ,"Control_temp","Treatment_temp") 
colnames(d2) <- Temp_name
Temp_name_sd <- c("Strain","Strain_number","Condition","Line","Generation","Highlight","Species","Line number","Control_temp_sd","Treatment_temp_sd") 
colnames(d3) <- Temp_name_sd


Control_temp_sd <- d3 %>% select(Control_temp_sd) %>% unlist()
Treatment_temp_sd <- d3 %>% select(Treatment_temp_sd) %>% unlist()
d2$Control_temp_sd <- Control_temp_sd
d2$Treatment_temp_sd <- Treatment_temp_sd
d2 <- d2 %>% mutate_if(is.numeric, ~round(., 4))
d2$Specie <- ordered(d2$Specie, levels = c("SE","SU","SA","SK","SJ","SM","SP","SC"))

d2 <- d2 %>% filter(Condition != "Ancestral")

d2_control <- d2 %>% filter (Condition == "Control")

d2_treatment <- d2 %>% filter (Condition == "Treatment")
```

```{r figure3B, fig.width=10, fig.height=6}
all4 <- d2 %>% ggplot(aes(x=Control_temp, y=Treatment_temp, col=Species, shape=Condition)) +
  geom_point(size=5, alpha=0.8)+
  geom_hline(yintercept = 0, linetype = 2)+
  geom_vline(xintercept = 0,linetype = 2)+
  xlim(c(-0.10,0.10))+
  ylim(c(-0.12,0.12))+
  scale_color_manual(values = c("#08519C","#88419D","#238B45","#2B8CBE","#737373","#980043","#D94801","#CB181D"),
                     labels = expression(italic("S. eubayanus"),italic("S. uvarum"),italic("S. arboricola"),italic("S. kudriavzevii"),italic("S. jurei"),italic("S. mikatae"),italic("S. paradoxus"),italic("S. cerevisiae")),
                     name= "Species") +
  scale_shape_manual(values = c(16,17),
                     name = "Evolution condition")+
  ylab("Relative fitness at final evolution temperature") +
  xlab("Relative fitness at 25°C") +
  annotate("text",x=-0.10, y=0.12,label="Increase at higher temp.",size=4,hjust = 0, vjust = 1)+
  annotate("text",x=-0.10, y=0.11,label="decrease at 25°C.",size=4,hjust = 0, vjust = 1)+
  annotate("text",x=0.10, y=-0.11,label="Increase at 25°C",size=4,hjust = 1, vjust = 0)+
  annotate("text",x=0.10, y=-0.12,label="decrease at higher temp.",size=4,hjust = 1, vjust = 0)+
  annotate("text",x=0.10, y=0.12,label="Increase at both temp.",size=4,hjust = 1, vjust = 1)+
  annotate("text",x=-0.10, y=-0.12,label="Decrease at both temp.",size=4,hjust = 0, vjust = 0)+
  theme_bw()+
  theme(axis.title.y = element_text( size = 14),
        axis.title.x = element_text( size = 14),
        axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14, angle=90,hjust = 0.5),
        legend.text = element_text(size = 14),legend.title = element_text(size = 14, face ="bold")) 


all4
ggsave(filename = file.path(output_dir, "Figure_3B.tiff"), plot = all4, width = 10, height = 6, dpi = 300)
```

Of the 64 I-populations, we observed 35 I-populations statistically significantly increasing in relative fitness at higher temperatures, of which 13 also decreased at 25 °C relative to their respective ancestral strains, while of the 64 C-populations, 6 C-populations significantly increased in relative fitness at 25 ºC. Of the 128 C and I-populations, four populations performed poorly after evolution at both constant and higher temperatures, and no populations showed a fitness gain in both constant and higher temperatures.

Linear mixed model:
The model explaining most of the variation in relative fitness included the interaction between temperature and evolution condition (C vs. I).
```{r}
reduced_model <- lmer(Rate_fitness ~ 1 + (1|Species/Highlight), data = d, REML = FALSE)
null_model1 <- lmer(Rate_fitness ~ Condition + (1|Species/Highlight), data = d, REML = FALSE)
null_model2 <- lmer(Rate_fitness ~ Temperature + (1|Species/Highlight), data = d, REML = FALSE)
full_model3 <- lmer(Rate_fitness ~ Temperature+Condition + (1|Species/Highlight), data = d, REML = FALSE)
full_model7 <- lmer(Rate_fitness ~ Temperature*Condition + (1|Species/Highlight), data = d, REML = FALSE)

anova(reduced_model, null_model1, null_model2,full_model3,full_model7)

full_model7 <- lmer(Rate_fitness ~ Temperature*Condition + (1|Species/Highlight), data = d)
summary(full_model7)
```

Populations evolved under increasing temperature (I-populations) reached a higher relative fitness by the end of the evolution experiment than the populations evolved under constant temperature (C-populations) when evaluated at the higher, strain-specific final evolution temperature.

Plotting figure 3C

```{r figure3C, fig.width=6, fig.height=6}
P4 <- d %>% 
  filter(Condition != "Ancestral") %>%
  ggplot(aes(x=Condition, y=Rate_fitness, fill=Condition)) +
  geom_hline(yintercept = 0, linetype = 2)+
  geom_boxplot(lwd=0.2) +
  scale_fill_manual(values = c("#4575B4","#D6604D")) +
  ylab("Relative fitness") +
  xlab("Evolution condition") +
  ggtitle("Growth Rate")+
  theme_bw()+
  theme(axis.title.y = element_text( size = 14),
        axis.title.x = element_text( size = 14),
        axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14),
        plot.title = element_text(size = 16,face = "bold",hjust = 0.5),
        legend.position = "none",
        strip.background = element_rect(colour="black", fill="white", linewidth=1, linetype="solid"),
        strip.text.x = element_text(size=12,face="bold"))+
  facet_grid(~Temperature,labeller = labeller(Temperature = temp.labs))
P4
ggsave(filename = file.path(output_dir, "Figure_3C.tiff"), plot = P4, width = 6, height = 6, dpi = 300)
```


We then evaluated relative fitness at the species level and tested for differences between cold-tolerant species (S. eubayanus, S. arboricola, S. kudriavzevii), warm-tolerant species (S. mikatae, S. paradoxus, S. cerevisiae), and species with an intermediate thermal profile (S. jurei, S. uvarum). We found differences in relative fitness between species after evolving populations under increasing temperature.

Plotting figure 3D
```{r figure3D, fig.width=6, fig.height=6}

temp.labs <- c("Higher temp.", "25°C")
names(temp.labs) <- c("Treatment", "Control")

P1 <- d %>% 
  filter(Condition != "Ancestral") %>%
  ggplot(aes(x=Species, y=Rate_fitness, fill=Condition)) +
  geom_hline(yintercept = 0, linetype = 2)+
  geom_boxplot(lwd=0.2) +
  scale_fill_manual(values = c("#4575B4","#D6604D")) +
  ylab("Relative fitness") +
  xlab("") +
  ggtitle("Growth Rate")+
  theme_bw()+
  guides(fill=guide_legend(title = "Evolution condition"))+
  theme(axis.title.y = element_text( size = 14),
        axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14),
        plot.title = element_text(size = 16,face = "bold",hjust = 0.5),
        legend.text = element_text(size = 14), legend.title = element_text(size = 16),
        legend.position = "bottom",
        strip.background = element_rect(colour="black", fill="white", linewidth=1, linetype="solid"),
        strip.text.x = element_text(size=12,face="bold"))+
  facet_grid(~Temperature,labeller = labeller(Temperature = temp.labs))

P1
ggsave(filename = file.path(output_dir, "Figure_3D.tiff"), plot = P1, width = 6, height = 6, dpi = 300)
```

Plotting figure S8

```{r figureS8, fig.width=12, fig.height=18}
SE <- d %>% 
  filter(Condition != "Ancestral", Species == "SE") %>%
  ggplot(aes(x=Strain_2, y=Rate_fitness, fill=Condition)) +
  geom_hline(yintercept = 0, linetype = 2)+
  geom_boxplot(lwd=0.2) +
  scale_fill_manual(values = c("#4575B4","#D6604D")) +
  ylab("Relative fitness") +
  xlab("Strain") +
  ggtitle("S. eubayanus")+
  theme_bw()+
  scale_y_continuous(limits = c(-0.1,0.12), breaks = seq(-0.1,0.12,0.04))+
  guides(fill=guide_legend(title = "Evolution condition"))+
  theme(axis.title.y = element_text( size = 14),
        axis.title.x = element_text( size = 14),
        axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14),
        plot.title = element_text(size = 16,face = "bold.italic",hjust = 0.5),
        legend.text = element_text(size = 14), legend.title = element_text(size = 16),
        strip.background = element_rect(colour="black", fill="white", linewidth=1, linetype="solid"),
        strip.text.x = element_text(size=12,face="bold"),
        strip.text.y = element_text(size=12,face="bold"))+
  facet_grid(~Temperature,labeller = labeller(Temperature = temp.labs))

SU <- d %>% 
  filter(Condition != "Ancestral", Species == "SU") %>%
  ggplot(aes(x=Strain_2, y=Rate_fitness, fill=Condition)) +
  geom_hline(yintercept = 0, linetype = 2)+
  geom_boxplot(lwd=0.2) +
  scale_fill_manual(values = c("#4575B4","#D6604D")) +
  ylab("Relative fitness") +
  xlab("Strain") +
  ggtitle("S. uvarum")+
  theme_bw()+
  scale_y_continuous(limits = c(-0.1,0.12), breaks = seq(-0.1,0.12,0.04))+
  guides(fill=guide_legend(title = "Evolution condition"))+
  theme(axis.title.y = element_text( size = 14),
        axis.title.x = element_text( size = 14),
        axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14),
        plot.title = element_text(size = 16,face = "bold.italic",hjust = 0.5),
        legend.text = element_text(size = 14), legend.title = element_text(size = 16),
        strip.background = element_rect(colour="black", fill="white", linewidth=1, linetype="solid"),
        strip.text.x = element_text(size=12,face="bold"),
        strip.text.y = element_text(size=12,face="bold"))+
  facet_grid(~Temperature,labeller = labeller(Temperature = temp.labs))

SA <- d %>% 
  filter(Condition != "Ancestral", Species == "SA") %>%
  ggplot(aes(x=Strain_2, y=Rate_fitness, fill=Condition)) +
  geom_hline(yintercept = 0, linetype = 2)+
  geom_boxplot(lwd=0.2) +
  scale_fill_manual(values = c("#4575B4","#D6604D")) +
  ylab("Relative fitness") +
  xlab("Strain") +
  ggtitle("S. arboricola")+
  theme_bw()+
  scale_y_continuous(limits = c(-0.1,0.12), breaks = seq(-0.1,0.12,0.04))+
  guides(fill=guide_legend(title = "Evolution condition"))+
  theme(axis.title.y = element_text( size = 14),
        axis.title.x = element_text( size = 14),
        axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14),
        plot.title = element_text(size = 16,face = "bold.italic",hjust = 0.5),
        legend.text = element_text(size = 14), legend.title = element_text(size = 16),
        strip.background = element_rect(colour="black", fill="white", linewidth=1, linetype="solid"),
        strip.text.x = element_text(size=12,face="bold"),
        strip.text.y = element_text(size=12,face="bold"))+
  facet_grid(~Temperature,labeller = labeller(Temperature = temp.labs))

SK <- d %>% 
  filter(Condition != "Ancestral", Species == "SK") %>%
  ggplot(aes(x=Strain_2, y=Rate_fitness, fill=Condition)) +
  geom_hline(yintercept = 0, linetype = 2)+
  geom_boxplot(lwd=0.2) +
  scale_fill_manual(values = c("#4575B4","#D6604D")) +
  ylab("Relative fitness") +
  xlab("Strain") +
  ggtitle("S. kudriavzevii")+
  theme_bw()+
  scale_y_continuous(limits = c(-0.1,0.12), breaks = seq(-0.1,0.12,0.04))+
  guides(fill=guide_legend(title = "Evolution condition"))+
  theme(axis.title.y = element_text( size = 14),
        axis.title.x = element_text( size = 14),
        axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14),
        plot.title = element_text(size = 16,face = "bold.italic",hjust = 0.5),
        legend.text = element_text(size = 14), legend.title = element_text(size = 16),
        strip.background = element_rect(colour="black", fill="white", linewidth=1, linetype="solid"),
        strip.text.x = element_text(size=12,face="bold"),
        strip.text.y = element_text(size=12,face="bold"))+
  facet_grid(~Temperature,labeller = labeller(Temperature = temp.labs))

SJ <- d %>% 
  filter(Condition != "Ancestral", Species == "SJ") %>%
  ggplot(aes(x=Strain_2, y=Rate_fitness, fill=Condition)) +
  geom_hline(yintercept = 0, linetype = 2)+
  geom_boxplot(lwd=0.2) +
  scale_fill_manual(values = c("#4575B4","#D6604D")) +
  ylab("Relative fitness") +
  xlab("Strain") +
  ggtitle("S. jurei")+
  theme_bw()+
  scale_y_continuous(limits = c(-0.1,0.12), breaks = seq(-0.1,0.12,0.04))+
  guides(fill=guide_legend(title = "Evolution condition"))+
  theme(axis.title.y = element_text( size = 14),
        axis.title.x = element_text( size = 14),
        axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14),
        plot.title = element_text(size = 16,face = "bold.italic",hjust = 0.5),
        legend.text = element_text(size = 14), legend.title = element_text(size = 16),
        strip.background = element_rect(colour="black", fill="white", linewidth=1, linetype="solid"),
        strip.text.x = element_text(size=12,face="bold"),
        strip.text.y = element_text(size=12,face="bold"))+
  facet_grid(~Temperature,labeller = labeller(Temperature = temp.labs))

SM <- d %>% 
  filter(Condition != "Ancestral", Species == "SM") %>%
  ggplot(aes(x=Strain_2, y=Rate_fitness, fill=Condition)) +
  geom_hline(yintercept = 0, linetype = 2)+
  geom_boxplot(lwd=0.2) +
  scale_fill_manual(values = c("#4575B4","#D6604D")) +
  ylab("Relative fitness") +
  xlab("Strain") +
  ggtitle("S. mikatae")+
  theme_bw()+
  scale_y_continuous(limits = c(-0.1,0.12), breaks = seq(-0.1,0.12,0.04))+
  guides(fill=guide_legend(title = "Evolution condition"))+
  theme(axis.title.y = element_text( size = 14),
        axis.title.x = element_text( size = 14),
        axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14),
        plot.title = element_text(size = 16,face = "bold.italic",hjust = 0.5),
        legend.text = element_text(size = 14), legend.title = element_text(size = 16),
        strip.background = element_rect(colour="black", fill="white", linewidth=1, linetype="solid"),
        strip.text.x = element_text(size=12,face="bold"),
        strip.text.y = element_text(size=12,face="bold"))+
  facet_grid(~Temperature,labeller = labeller(Temperature = temp.labs))

SP <- d %>% 
  filter(Condition != "Ancestral", Species == "SP") %>%
  ggplot(aes(x=Strain_2, y=Rate_fitness, fill=Condition)) +
  geom_hline(yintercept = 0, linetype = 2)+
  geom_boxplot(lwd=0.2) +
  scale_fill_manual(values = c("#4575B4","#D6604D")) +
  ylab("Relative fitness") +
  xlab("Strain") +
  ggtitle("S. paradoxus")+
  theme_bw()+
  scale_y_continuous(limits = c(-0.1,0.12), breaks = seq(-0.1,0.12,0.04))+
  guides(fill=guide_legend(title = "Evolution condition"))+
  theme(axis.title.y = element_text( size = 14),
        axis.title.x = element_text( size = 14),
        axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14),
        plot.title = element_text(size = 16,face = "bold.italic",hjust = 0.5),
        legend.text = element_text(size = 14), legend.title = element_text(size = 16),
        strip.background = element_rect(colour="black", fill="white", linewidth=1, linetype="solid"),
        strip.text.x = element_text(size=12,face="bold"),
        strip.text.y = element_text(size=12,face="bold"))+
  facet_grid(~Temperature,labeller = labeller(Temperature = temp.labs))

SC <- d %>% 
  filter(Condition != "Ancestral", Species == "SC") %>%
  ggplot(aes(x=Strain_2, y=Rate_fitness, fill=Condition)) +
  geom_hline(yintercept = 0, linetype = 2)+
  geom_boxplot(lwd=0.2) +
  scale_fill_manual(values = c("#4575B4","#D6604D")) +
  ylab("Relative fitness") +
  xlab("Strain") +
  ggtitle("S. cerevisiae")+
  theme_bw()+
  scale_y_continuous(limits = c(-0.1,0.12), breaks = seq(-0.1,0.12,0.04))+
  guides(fill=guide_legend(title = "Evolution condition"))+
  theme(axis.title.y = element_text( size = 14),
        axis.title.x = element_text( size = 14),
        axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14),
        plot.title = element_text(size = 16,face = "bold.italic",hjust = 0.5),
        legend.text = element_text(size = 14), legend.title = element_text(size = 16),
        strip.background = element_rect(colour="black", fill="white", linewidth=1, linetype="solid"),
        strip.text.x = element_text(size=12,face="bold"),
        strip.text.y = element_text(size=12,face="bold"))+
  facet_grid(~Temperature,labeller = labeller(Temperature = temp.labs))

Curve1 <- ggarrange(SE,SU,SA,SK,SJ,SM,SP,SC, 
                    labels = c("A", "B","C","D","E","F","G","H"),
                    ncol = 2, nrow = 4, common.legend = TRUE, legend = "bottom")

figureS8 <- annotate_figure(Curve1,
                top = text_grob("Fitness", color = "black", face = "bold", size = 16))
figureS8
ggsave(filename = file.path(output_dir, "Figure_S8.tiff"), plot = figureS8, width = 12, height = 18, dpi = 300)
```

Plotting figure 3

```{r figure3, fig.width=12, fig.height=18}
figure3 <- ggdraw() +
  draw_plot(ev2, 0.05, .7, 0.9, .3) +
  draw_plot(all4, 0.05, 0.35, .9, .35) +
  draw_plot(P4, 0, 0.05, .45, .30) +
  draw_plot(P1, 0.45, 0, .55, .35) +
  draw_plot_label(c("A", "B", "C","D"), c(0, 0, 0, 0.45), c(1, 0.7, 0.35, 0.35), size = 15)
figure3

ggsave(filename = file.path(output_dir, "Figure_3.tiff"), plot = figure3, width = 12, height = 18, dpi = 300)
```


On average, I-populations of *S. eubayanus*, *S. uvarum*, *S. arboricola*, *S. kudriavzevii*, *S. jurei*, and *S. mikatae* significantly improved in relative fitness at higher temperatures, but populations of the more warm-tolerant species *S. paradoxus* and *S. cerevisiae* did not. 


  
  
  
