---
title: "Section: *Saccharomyces* species differ in thermal performance"
author: "Jennifer Molinet"
date: "2025-01-17"
output: html_document
---

We first determined the natural genetic diversity in temperature tolerance across eight species of the *Saccharomyces* genus at ten different temperatures under microculture conditions. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Libraries**

```{r}
suppressMessages(library(tidyverse))
library(ggplot2)
library("FactoMineR")
suppressMessages(library("factoextra"))
library(reshape2)
library(lme4)
```

**Directories**

```{r}
output_dir <- "../results/Section1/"
input_dir <- "../data/Section1"
```

**Loading data**
```{r}
file_path <- file.path(input_dir, "Kinetic_parameters.txt")
d <- read.table(file_path, header = TRUE, stringsAsFactors = FALSE)
head(d)
```

**Analyzing growth rates at different temperatures**

```{r}
d$Specie <- ordered(d$Specie, levels = c("SE","SU","SA","SK","SJ","SM","SP","SC"))
d$Temperature <- as.character(d$Temperature)
```

We first determined the natural genetic diversity in temperature tolerance across eight species of the Saccharomyces genus at ten different temperatures under microculture conditions. We evaluated 51 genetically distinct strains across eight species, representative of their population structure, different geographic areas, and ecological niches (Figure S2). 

Plotting figure S2: Maximum growth rate at different temperatures for Saccharomyces species.

```{r figureS2, fig.width=15, fig.height=6}
figureS2 <- d %>% 
  ggplot(aes(x=Temperature, y=Rate, fill=Specie)) +
  geom_dotplot(binaxis='y', stackdir='center', dotsize=0.5) +
  scale_fill_manual(values = c("#08519C","#88419D","#238B45","#2B8CBE","#737373","#980043","#D94801","#CB181D"),
                    labels=c("SE" = "S. eubayanus", "SU" = "S. uvarum", "SA" = "S. arboricola",
                             "SK" = "S. kudriavzevii", "SJ" = "S. jurei", "SM" = "S. mikatae", "SP"="S. paradoxus","SC"="S. cerevesiae"),
                    name = "Species") +
  ylab("umax (OD/h)") +
  xlab("Temperature (ºC)") +
  ggtitle("Growth rate at different temperatures") +
  theme_bw() +
  scale_y_continuous(limits = c(0.0,1), breaks = seq(0.0,1,0.2))+
  theme(axis.title.y = element_text( size = 14, face = "bold"),
        axis.title.x = element_text( size = 14, face = "bold"),
        axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14),
        plot.title = element_text(size = 16,face = "bold",hjust = 0.5),
        legend.text = element_text(size = 16), legend.title = element_text(size = 16, face = "bold")) 
figureS2

ggsave(filename = file.path(output_dir, "Figure_S2.tiff"), plot = figureS2, width = 15, height = 6, dpi = 300)
```

Growth rates at higher temperatures (28-40 °C) showed a larger dispersion of variation between species compared to growth rates at lower temperatures (16-23 °C) 

We performed a principal component analysis (PCA) to describe the phenotypic landscape across all 51 strains before experimental evolution (Figure 2A).

```{r}
d_ave <- d %>% 
  group_by(Strain,Specie,Temperature) %>%
  dplyr::summarise(., sd = sd(Rate),
                   ave_rate = mean(Rate)) %>%
  ungroup()
head(d_ave)
```

```{r}
d2 <- d_ave %>% select(Strain,Specie,ave_rate, Temperature) %>%
  dcast(Strain+Specie~Temperature, value.var = "ave_rate")
head(d2)
```

```{r}
d3 <- d2 %>% select(-Strain, -Specie)
head(d3)
res.pca <- PCA(d3, graph = FALSE)
```

Plotting figure 2A: PCA using the maximum specific growth rate at ten different temperatures.

```{r figure2A, fig.width=12, fig.height=12}
pca <- fviz_pca_biplot(res.pca, 
                # Fill individuals by groups
                geom.ind = "point",
                pointshape = 21,
                pointsize = 2.5,
                fill.ind = d2$Specie,
                col.ind = "black",
                # Color variable by groups
                col.var = factor(c("16°C","18°C","20°C","23°C","25°C","28°C","31°C","34°C","37°C","40°C")),
                xlab = "PC1 (44.1%)", ylab = "PC2 (34.3%)",
                legend.title = list(fill = "Species", color = "Temperature"),
                repel = TRUE,        # Avoid label overplotting
                mean.point = FALSE,
                addEllipses = TRUE, ellipse.type = "convex"
)+
  ggpubr::fill_palette(c("#08519C","#88419D","#238B45","#2B8CBE","#737373","#980043","#D94801","#CB181D"))+      # Indiviual fill color
  ggpubr::color_palette(c("#0000f9","#2121ff","#9797ff","#b1b1ff","#ffa500","#e65b00","#ff7f7f","#ff3232","#e50000","#990000"))+      # Variable colors
   ggtitle("PCA")+
  theme(axis.title.y = element_text( size = 14, face = "bold"),
        axis.title.x = element_text( size = 14, face = "bold"),
        axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14),
        plot.title = element_text(size = 16,face = "bold",hjust = 0.5),
        legend.text = element_text(size = 16), legend.title = element_text(size = 16, face = "bold")) 


pca

ggsave(filename = file.path(output_dir, "Figure_2A.tiff"), plot = pca, width = 12, height = 12, dpi = 300)
```

PC2 separates two temperature groups: colder temperatures between 18 and 23 °C and hotter temperatures between 28 and 40 °C. PC1 separates, to some degree, the eight species. The individual factor map shows that strains of S. cerevisiae, S. paradoxus, and S. mikatae fall into the right lower quarter of phenotype space and are associated with higher growth rates at higher temperatures. 

Hierarchical clustering was performed to confirm differences observed between species.

```{r}
row.names(d3) <- d2$Strain
res.pca <- PCA(d3,ncp = 3, graph = FALSE)
```

```{r}
res.hcpc2 <- HCPC(res.pca, nb.clust = 3, consol = FALSE) 
cluster_colors <- c( "#2E9FDF","#FC4E07", "#E7B800")
```

Plottinf figure S3: Hierarchical clustering of Saccharomyces strains.

```{r figure3A, fig.width=12, fig.height=12}
cluster3 <- fviz_dend(res.hcpc2, k = 3, # Cut in four groups
                      cex = 0.6, # label size
                      k_colors = c("#2E9FDF", "#E7B800", "#FC4E07"),
                      color_labels_by_k = TRUE, # color labels by groups
                      rect = TRUE # Add rectangle around groups
)+
    ggtitle("Cluster Dendrogram")+
   theme(axis.title.y = element_text( size = 14, face = "bold"),
        axis.text.y = element_text(size = 14),
        plot.title = element_text(size = 16,face = "bold",hjust = 0.5)) 


cluster3

ggsave(filename = file.path(output_dir, "Figure_S3A.tiff"), plot = cluster3, width = 12, height = 12, dpi = 300)
```

```{r figure3B, fig.width=12, fig.height=12}
cluster2 <- fviz_cluster(res.hcpc2,
             repel = TRUE,            # Avoid label overlapping
             show.clust.cent = FALSE, # Show cluster centers
             palette = cluster_colors,         # Color palette see ?ggpubr::ggpar
             ggtheme = theme_minimal(),
             xlab = "PC1 (44.1%)", ylab = "PC2 (34.3%)",
             main = "Factor map"
)+
  ggtitle("Individual Factor Map")+
  theme(axis.title.y = element_text( size = 14, face = "bold"),
        axis.title.x = element_text( size = 14, face = "bold"),
        axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14),
        plot.title = element_text(size = 16,face = "bold",hjust = 0.5),
        legend.text = element_text(size = 16), legend.title = element_text(size = 16, face = "bold")) 

cluster2

ggsave(filename = file.path(output_dir, "Figure_S3B.tiff"), plot = cluster2, width = 12, height = 12, dpi = 300)
```
We observed three main clusters with most strains of S. cerevisiae, S. paradoxus, and S. mikatae grouping together in one cluster.

Linear mixed models indicated significant effect of temperature on growth rate and interspecies variability, which was more important in explaining differences in growth rate than between strains within species.

```{r}
anova_model <- lmer(Rate ~ Temperature + (1 | Specie/Strain), data = d,REML = FALSE)
anova_model_no_random <- lm(Rate ~ Temperature, data = d)
anova_model_random <- lmer(Rate ~ 1 + (1 | Specie/Strain), data = d, REML = FALSE)
anova(anova_model, anova_model_no_random, anova_model_random)

anova_model <- lmer(Rate ~ Temperature + (1 | Specie/Strain), data = d)
summary(anova_model)
```

```{r}
null_model3 <- lmer(Rate ~ Temperature + (1|Specie/Strain), data = d)
summary(null_model3)
```

```{r}
var_comp <- as.data.frame(VarCorr(null_model4))

var_total <- sum(var_comp$vcov)
prop_var_especie <- var_comp$vcov[2] / var_total
prop_var_cepa <- var_comp$vcov[1] / var_total

print(paste("Variance ratio between species:", prop_var_especie))
print(paste("Proportion of variance between strains within species:", prop_var_cepa))
```

